---
title: "Introduction to genieR"
author: "Fei Xiang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette demonstrates the use of the package. An HIV phologenetic tree is included in the package directory. To visualise the tree:

```{r message=FALSE}
library(genieR)
```

```{r}
library(ape)
data(vill1_lsd)
plot(vill1_lsd)
```

Then we can fit a exponential growth coalescent model for this tree by
```{r}
fit1=Geniefit(vill1_lsd,Model="expo",start=c(100,.1),upper=Inf,lower=0)
```


This package can also simulates coalescent times for isochronous/heterochronous data. 
```{r}
 trajectory<-function(x)  exp(10*x)
 sample1<-cbind(c(9,1,2,1),c(0,.008,.03,.1))
 example_hetero<-coalgen_hetero(sample1, trajectory)
 sample<-c(100,0)
 example_iso<-coalgen_iso(sample, trajectory)
```

This package can also extracts coalescent and sampling times from a given phylogeny.
```{r}
library(ape)
 t1=rcoal(20)
heterochronous.gp.stat(t1)
```



